---
title: "linear model from Web"
author: "Desmond Koutso"
date: "Summer 2022"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot

```{r}
plot(SIMS~ARM,data=data)
lmr = lm(SIMS~ARM,data=data)
abline(lmr)
```

From this scatter plot of Model 1 SIM~ARMS have a positve association as the line passes through the middle of the plots.  


### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```

The Numerical results gives 0.6860073.  

### Inferential  (Build model.1)

```{r}
summary(lmr)
newdata=c(GRIP=94, ARM=88)
```

From the inferential results, the P-value is 0.00000000000000002.  
The adjusted R-squared is 0.467, and the multiple R-squared is 0.4706.  

#### Predict at target point

```{r}
predictSIM = predict.lm(lmr, newdata=data.frame(GRIP = 94, ARM = 88))
print(predictSIM)
```

The predicted target piont is 0.7063836.  

#### scatterplot with model fit

```{r}
plot(lmr)
```

From the plot residuals vs leverage, there is a great concentration of the number of plot on the leverage axis 0.01 and standardized residuals axis on -1 till 2.      

## Model 2 SIM~GRIP

### Now add in scatterplot

```{r}
plot(SIMS~GRIP,data=data)
gp = lm(SIMS~GRIP,data=data)
abline(gp)
```

From this scatter plot Model 2 SIM~GRIP, there is a line that passes through the middle of the plots, which implies there is a positive association between the variables SIMS and GRIP.   


### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

The numerical results gives 0.6398458.  

### Inferential  (Build model.2)

```{r}
summary(gp)
newdata=c(GRIP=94, ARM=88)
```
  
From the inferential results, the p-value is 0.00000000000000002.  
The multiple R-squared gives 0.4094, and the adjusted R-squared gives 0.4053.  
  

#### predict model.2 at target point

```{r}
predictSIM = predict.lm(gp, newdata=data.frame(GRIP = 94, ARM = 88))
print(predictSIM)
```
The prediction for the model.2 at target point is -0.5361543.  

#### now add the model fit to our plot for model.2
  
```{r}
plot(gp)
```

Based on the scattered plot residuals vs fitted, there is a curve that passes through the plotts. This implies that there is a variable that does not increase at a certain level.  

## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

```{r}
data2=lm(SIMS~ARM+GRIP,data=data)
summary(data2)
```

From the numerical results, the p-value is 0.00000000000000000.  
The multiple R-squared gives 0.5422, and the Adjusted R-squared is 0.5358.  
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point

```{r}
predict.lm(data2, newdata=data.frame(GRIP=94, ARM=88))
```
 The prediction of model.3 at target point is 0.1496476.  
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(lmr, data2)
```

Comparing the Model.1 vs Model.3 using annova test, The sum of sq gives 29.45.  

### Model.2 vs Model.3

```{r}
anova(gp, data2)
```

Comparing the Model.1 vs Model.3 using annova test, The sum of sq gives 54.639.    


## Informally compare Model.1 with model.2

Comparing model 1 with model 2, the residual difference in model1 is 145, slightly higher than that of model 2 which is 144.  
The R-squared for model1 is 243.07, slightly higher than model2, which is 188.43.  